\section{Introduction}
\label{sec:intro}

\subsection{Free-energy computation along a reaction coordinate}
A central task in computational statistical mechanics is to compute the free
energy associated with a low-dimensional \emph{reaction coordinate} of a
high-dimensional system in thermal equilibrium~\citep{LelievreRoussetStoltz2010}.
Let $q=(x,y)\in\R^2$ denote a configuration\footnote{We write a configuration
as $q=(x,y)$. The symbol $q_t(x)$, always carrying a time subscript and a scalar
argument, denotes instead the Fisher--Rao \emph{target marginal} of
\cref{sec:fr}; the two are never used in the same role.} drawn from the
canonical (Boltzmann--Gibbs) density
\begin{equation}
  \pi(x,y) \;\propto\; e^{-\beta V(x,y)},
  \label{eq:canonical}
\end{equation}
where $V$ is the potential energy and $\beta=1/(k_{\mathrm B}T)$ is the inverse
temperature. Fixing the reaction coordinate to the first coordinate,
$\xib(x,y)=x$, the free energy $F(x)$ and the (local) mean force $F'(x)$ along
$\xib$ are the objects of interest. The difficulty is metastability: when $V$
has well-separated basins along $x$, unbiased Langevin dynamics crosses the
barrier rarely, so a direct estimate of $F$ converges intolerably slowly.

\subsection{Adaptive Biasing Force}
The Adaptive Biasing Force (ABF) method~\citep{DarvePohorille2001,Henin2004,
Comer2015} removes this metastability adaptively. It maintains a running
estimate $\Fphat_t(x)$ of the mean force and applies the antiderivative
$\Bt(x)=\int^x\Fphat_t$ as a bias that depends only on $x$. As
$\Bt\to F$, the effective dynamics along $x$ becomes diffusive and the
reaction coordinate is explored freely, while --- crucially --- the bias does
not change the conditional law of the orthogonal degrees of freedom given $x$.
ABF is mathematically well understood, with long-time convergence
guarantees~\citep{LelievreRoussetStoltz2008}, and is a natural baseline against
which to measure any proposed enhancement.

\subsection{Why add a Fisher--Rao correction?}
ABF accelerates exploration through the \emph{force}, but it does not directly
control \emph{where the particles are}. Early in a simulation the estimated bias
is poor, so the empirical reaction-coordinate marginal $\phat_t(x)$ can remain
concentrated in the initial basin, starving other values of $x$ of the samples
ABF needs to estimate $F'(x)$ there. A Fisher--Rao (FR) birth--death
correction~\citep{LuLuNolen2019} addresses exactly this: interpreting the
evolution of $\phat_t$ as a gradient flow of relative entropy in the
Fisher--Rao geometry~\citep{Amari2016}, one reallocates probability mass by
\emph{killing} particles in over-represented regions and \emph{cloning}
particles in under-represented regions, driving $\phat_t$ toward a chosen target
marginal $q_t(x)$ without integrating any additional dynamics. Coupling such a
correction to ABF is appealing because the two act on complementary objects
(force versus density), but it is not obvious that it helps: a birth--death step
resamples \emph{whole} configurations $(x,y)$, and if it distorts the
conditional law $Y\mid X=x$ it will corrupt precisely the quantity ABF
estimates.

\subsection{Central question and contributions}
This report asks three questions on a controlled example where the exact answer
$\Fref$ is available by quadrature:
\begin{enumerate}
  \item \textbf{Does FR help ABF?} (Q1)
  \item \textbf{How should FR be added so that it helps?} (Q2--Q3)
  \item \textbf{What are the limitations, failure modes, and future
        directions?} (Q4)
\end{enumerate}
Our contributions are: (i) a precise statement of the coupled ABF--FR algorithm
in consistent notation (\cref{sec:abf,sec:fr}); (ii) a large production study
(\NRuns{} runs) that quantifies, with matched seeds, the effect of FR and of its
target marginal (\cref{sec:design,sec:results}); and (iii) a mechanistic
analysis (\cref{sec:mechanism}) separating the contribution of the FR
\emph{mechanism} from that of the FR \emph{target}. A recurring methodological
point is the careful interpretation of an \emph{oracle} target that uses
$\Fref$: it is a diagnostic control that bounds the achievable benefit, never a
deployable method. We find that the practical, estimated-target FR is a
\emph{modest convergence accelerator} here, and we delineate the conditions
under which it helps and those under which it could hurt.
