\section{The Fisher--Rao marginal correction}
\label{sec:fr}

The Fisher--Rao (FR) correction is a birth--death / particle-reallocation step
that acts on the \emph{reaction-coordinate marginal} of the ensemble. It is
interleaved with the ABF dynamics~\eqref{eq:abf_sde} and is designed to drive the
empirical marginal $\phat_t(x)$ toward a target marginal $q_t(x)$, without
adding any drift to the $(x,y)$ dynamics.

\subsection{Empirical and target marginals}
At step $n$, after the Langevin proposal of all $N$ replicas, the empirical
reaction-coordinate marginal is estimated by a (reflected, edge-corrected)
kernel density / binned-smoothed estimate of bandwidth $\eta$,
\begin{equation}
  \phat_{n}(x) \;=\; \frac{1}{N}\sum_{i=1}^{N} K_\eta\!\bigl(x-X_n^i\bigr),
  \label{eq:phat}
\end{equation}
renormalised on the $x$-grid. The target $q_n(x)$ is one of three choices
(\cref{sec:targets}); all are nonnegative and normalised, $\int q_n=1$.

\subsection{Fisher--Rao score}
For each particle the FR \emph{score} is the pointwise log-ratio of the
empirical to the target marginal, with its ensemble mean removed:
\begin{equation}
  S_i \;=\;
  \log\frac{\phat_n(X^i_n)+\varepsilon}{q_n(X^i_n)+\varepsilon}
  \;-\;
  \underbrace{\int \phat_n(z)\,\log\frac{\phat_n(z)}{q_n(z)}\dd z}
  _{=\,\mathrm{KL}(\phat_n\,\Vert\,q_n)} ,
  \label{eq:score}
\end{equation}
where $\varepsilon>0$ is a small regulariser and the integral baseline equals
the Kullback--Leibler divergence $\mathrm{KL}(\phat_n\Vert q_n)$ (a grid
quadrature; it coincides with the Monte-Carlo baseline
$\tfrac1N\sum_k\log[\phat_n(X^k)/q_n(X^k)]$ up to discretisation). Subtracting
the baseline makes the score mean-zero, so the birth--death step below conserves
the expected particle number. The score is clipped to $[-S_{\max},S_{\max}]$
for stability. Its interpretation is direct:
\begin{itemize}
  \item $S_i>0$: $X^i$ lies where $\phat_n>q_n$, an \emph{over-represented}
        region (too many particles);
  \item $S_i<0$: $X^i$ lies where $\phat_n<q_n$, an \emph{under-represented}
        region (too few particles).
\end{itemize}

\subsection{Birth--death resampling}
The FR step realises the Fisher--Rao gradient flow of
$\mathrm{KL}(\phat\Vert q)$ as a fixed-$N$ birth--death process. Over a sub-step
of length $\Delta t$ and at effective rate $\gamma_n^{\mathrm{eff}}$
(\cref{sec:schedule}), each \emph{over-represented} particle ($S_i>0$) is killed
with probability
\begin{equation}
  \mathbb P(\text{$i$ dies}) \;=\; 1-\exp\!\bigl(-\gamma_n^{\mathrm{eff}}\,S_i\,\Delta t\bigr),
  \label{eq:death}
\end{equation}
and every killed slot is overwritten by a \emph{clone} of an
\emph{under-represented} particle, sampled from the pool $\{i:S_i<0\}$ with
weights $\propto|S_i|$. Cloning copies the \emph{whole} configuration
$(X^i,Y^i)$ --- the score depends only on $x$, but the $y$-coordinate must travel
with its $x$-coordinate, or \cref{prop:cond_invariance} would be violated. The
step is made deliberately gentle by three safeguards: (i) a burn-in and soft
ramp on the rate; (ii) score clipping; and (iii) a hard cap
$f_{\max}$ on the fraction of particles replaced per event
(\texttt{max\_event\_fraction}), enforced by randomly sub-sampling deaths when
the cap is exceeded. Net effect: gently replace particles in crowded regions of
$x$ by copies from starved regions, leaving $N$ fixed.

\subsection{Rate schedule}
\label{sec:schedule}
The FR rate is zero during an initial burn-in of $n_{\mathrm{burnin}}=
\lfloor b\,n_{\mathrm{steps}}\rfloor$ steps (burn-in fraction $b$), then ramps
softly to its maximum $\gamma$,
\begin{equation}
  \gamma_n^{\mathrm{eff}} \;=\;
  \gamma\,\Bigl(1-e^{-(n-n_{\mathrm{burnin}})/n_{\mathrm{ramp}}}\Bigr),
  \qquad n>n_{\mathrm{burnin}},
  \label{eq:ramp}
\end{equation}
and the birth--death step is applied every $\texttt{fr\_every}$ steps. The
burn-in lets ABF build a usable mean-force estimate (hence a usable target)
before any particles are moved; the ramp avoids a sudden shock to the ensemble.

\subsection{Target marginals: estimated, uniform, oracle}
\label{sec:targets}
The three targets differ only in which density $q_n$ the score~\eqref{eq:score}
is taken against. Recall $\Bt=\Fhat_n$ is the current ABF bias.

\paragraph{Estimated target (the practical method).}
A target free energy $\Fhat_n^{\mathrm{target}}$ is maintained independently of
the bias by an exponential moving average of the ABF estimate,
$\Fhat_{n+1}^{\mathrm{target}}=(1-r)\,\Fhat_n^{\mathrm{target}}+r\,\Fhat_{n+1}$,
and the target marginal is the \emph{biased} marginal that this target free
energy would induce under the current bias,
\begin{equation}
  q_n^{\mathrm{est}}(x) \;\propto\;
  \exp\!\Bigl[-\beta\bigl(\Fhat_n^{\mathrm{target}}(x)-\Bt(x)\bigr)\Bigr].
  \label{eq:q_est}
\end{equation}
This uses only quantities available at run time and is the method whose
practical value we assess.

\paragraph{Uniform target (a naive ablation).}
\begin{equation}
  q_n^{\mathrm{unif}}(x) \;=\; \mathrm{Unif}(x)\quad\text{on the $x$-grid.}
  \label{eq:q_unif}
\end{equation}
This tests whether simply \emph{flattening} the reaction-coordinate marginal is
enough. It is a target ablation, not the intended method; it ignores the fact
that, under a bias $\Bt\neq F$, the canonical biased marginal is not uniform.

\paragraph{Oracle target (a diagnostic control only).}
\begin{equation}
  q_n^{\mathrm{oracle}}(x) \;\propto\;
  \exp\!\Bigl[-\beta\bigl(\Fref(x)-\Bt(x)\bigr)\Bigr].
  \label{eq:q_oracle}
\end{equation}
\begin{remark}[The oracle target is not an algorithm]
\label{rem:oracle}
Equation~\eqref{eq:q_oracle} replaces the estimated target free energy by the
\emph{reference} free energy $\Fref$ --- which is exactly the unknown quantity
the whole procedure is trying to compute. The oracle target therefore
\emph{cannot} be evaluated in any real free-energy calculation. We include it
solely as a \emph{diagnostic / positive / ideal-information control}: it
measures how much the FR mechanism \emph{could} help if the target marginal
direction were specified perfectly. Throughout this report the oracle is read
only in that way; we never claim that ``oracle FR is the best method'' or that
``one should use the oracle target''. The practical comparison is always
\[
  \text{ABF only}\qquad\text{versus}\qquad\text{ABF}+\text{estimated-target FR},
\]
and the estimated-versus-oracle gap is interpreted as the headroom available
from improving the online target estimate.
\end{remark}

\subsection{The coupled ABF--FR step}
One outer iteration, given the state $\{(X^i_n,Y^i_n)\}$ and accumulators, is:
\begin{enumerate}
  \item update the ABF accumulators from the current particles and recompute
        $\Fphat_n,\Fhat_n=\Bt$ (\cref{sec:abf_estimator});
  \item propose the biased Langevin move~\eqref{eq:abf_sde} for all replicas;
  \item if FR is active at this step (past burn-in, on the \texttt{fr\_every}
        schedule): form $\phat_n$~\eqref{eq:phat} and the target $q_n$ ---
        one of~\eqref{eq:q_est},~\eqref{eq:q_unif},~\eqref{eq:q_oracle} ---
        compute the scores~\eqref{eq:score}, and apply the birth--death
        step~\eqref{eq:death};
  \item record diagnostics at the evaluation cadence.
\end{enumerate}
The ABF estimator sees the post-resampling ensemble at the next step, which is
the route by which improved $x$-coverage feeds back into a better mean-force
estimate. The remainder of the report quantifies whether, and how, this feedback
is beneficial.
