\section{Discussion, limitations, and future directions}
\label{sec:discussion}

\subsection{What the three cases establish (Q4)}
The study establishes a regime map rather than a universal ranking. FR is a mild
accelerator, and sometimes a source of extra noise, when ABF is already
well-sampled. It becomes valuable when ABF is starved of reaction-coordinate
support. In the many-body WCA case the same mechanism has a sharper operational
limit: too much birth--death destroys ancestor diversity faster than the dense
solvent can decorrelate cloned configurations. The different conclusions across
examples are therefore not a defect of the study; they are the main information
returned by the three-case design.

\subsection{Limitations}
We remain deliberately conservative about scope.
\begin{enumerate}
  \item \textbf{The reaction coordinates are low-dimensional.} All three systems
        use a one-dimensional reaction coordinate (a Cartesian component in the
        2D cases, a scalar bond length in WCA). Multi-dimensional or strongly
        curved collective variables, where the geometric mean-force term and the
        conditional sampling are far harder, are untested.
  \item \textbf{The oracle target is not a method or an upper bound.} In every
        case the oracle uses the reference free energy, the unknown being
        computed, and is reported only as a diagnostic control (\cref{rem:oracle}).
        Its behaviour is read comparatively: best in the easy metastability
        regime, tied with practical targets in WCA, and worse than ABF in the cold
        entropic bottleneck.
  \item \textbf{Production estimators are validated surrogates.} The 2D GPU
        backend uses a binned-smoothed mean-force/marginal estimator and a
        vectorised fixed-$N$ birth--death variant; the WCA backend uses a
        constrained projection with the geometric term and a direct
        death-replacement step. These pass their validation suites and agree with
        the kernel/TI references to within tolerance, but are not bit-identical to
        the original kernel notebook implementation.
  \item \textbf{Conditional diagnostics are incomplete in the many-body case.}
        The entropic-bottleneck case gives an analytic conditional check and shows
        no systematic collapse, but FR is not uniformly better than ABF in every
        bin. The WCA conditional fidelity is argued structurally
        (\cref{rem:manybody}) rather than measured per configuration.
  \item \textbf{Entropic-specific gains are real but target-limited.} In the base
        Case~II parameterisation the energetic double well dominates the difficulty
        and the $\omega_{\mathrm{in}}$ sweep is honestly flat. The dedicated
        entropy-dominant stress test (\cref{sec:eb_entropy_dominant}), which holds
        the \emph{total} barrier fixed and grows the entropic share, sharpens this:
        the \emph{deployable} gain still does not scale with the entropic fraction
        (slope $\EDBestSlope\%$ per unit $\phi$), whereas the \emph{oracle} gain
        does (slope $\EDBoracleSlope\%$, up to $+\EDBoraclePeak\%$). The
        entropic-specific benefit exists but is gated by the quality of the online
        free-energy target; the practical estimator does not realise it here.
\end{enumerate}

\subsection{Failure modes}
The analysis points to two related failure modes. First, an \emph{overly
aggressive mechanism} erodes ancestor diversity faster than the orthogonal
coordinates can decorrelate clones. This is realised in the WCA dimer for
\texttt{fr\_rate}$\ge1$ in the rate ladder, with even \texttt{fr\_rate}$=0.5$
turning harmful at production budget, and is masked in the entropic bottleneck
because its analytic channel re-equilibrates rapidly. Second, a \emph{misaligned
target} can hurt when it over-commits to a fixed shape during a transient ABF
bias. The metastability oracle--estimated gap quantifies target-direction
headroom only in the easy regime; the entropic-bottleneck oracle shows that this
reading does not generalise.

\subsection{Future directions}
\begin{itemize}
  \item \textbf{Adaptive, diversity-aware FR strength.} Tie \texttt{fr\_rate}/
        $\gamma$ to measured ancestor ESS or orthogonal-relaxation diagnostics, so
        the rate self-limits before diversity collapse.
  \item \textbf{Higher-dimensional and curved reaction coordinates}, where
        conditional sampling is genuinely hard and the geometric term is
        nontrivial.
  \item \textbf{Configuration-resolved conditional diagnostics} for the many-body
        case, to measure conditional fidelity rather than infer it from structure.
  \item \textbf{A better online target for entropic bottlenecks.} The
        entropy-dominant stress test (\cref{sec:eb_entropy_dominant}) shows the
        achievable (oracle) gain is entropic-specific but the ABF-bias estimate is
        too noisy to realise it. A lower-variance online free-energy estimator
        inside the FR target (e.g.\ a slower-EMA or eABF/CZAR-style estimate) is the
        natural way to convert that diagnosed headroom into a deployable gain.
\end{itemize}
